FACULTY EVALUATION ANALYSIS USING LIKERT SCALE
========================================================
The data for each department may be stored separately in files
and complete picture of each department can be had in a pdf 
file with groupings of faculty member names...

library(likert)
library(reshape)

First we transpose the data from column 9 to 109 

```{r}
al <- read.csv('All Proforma Data (Faculty Survey).csv',header=T)
resp <- t(al[,9:108])
colnames(resp) <- t(al[,8]) 

```

We trim unused rows at the last filled with the NAs and add faculty member names so many times as many are observations for each
```{r}
resp<- resp[rowSums(resp,na.rm=T)!=0,]
obs<-c()
for (i in seq(1,ncol(resp),18)) 
  obs <- append(obs,sum(rowSums(resp[,i:c(i+17)],na.rm=T)!=0))

#splitting each succesive 18 columns and binding them  down in to rows
totalobs = max(obs*length(obs))
maxobs = max(obs)
final<- c(NULL)
i=1
for (j in seq(1, totalobs, maxobs) )
  {
  final <- rbind(final, resp[1:maxobs,i:c(i+17)])
  i=i+18
}

## adding faculty member names as row names
row.names(final) <- rep(unique(al[,5]), each=maxobs)

## cleaning rows with full NAs
final <- final [rowSums(final,na.rm=T)!=0,]
write.csv(final, 'FacEvalforlikert.csv')



```

Upto here we get clean data if we also want to include department names then we could add a column in the beggining and add just like we added faculty member names.

Next we replace values 1 to 5 with the character equivalents and converting them into factors... 
```{r}
final <- as.data.frame(final)
final[final==5]<-'Strongly Agree'
final[final==4]<-'Agree'
final[final==3] <- 'Somewhat Agree'
final[final==2] <- 'Disagree'
final[final==1] <- 'Strongly Disagree'
colnames(final) <- as.character(al[1:18,7])
for (i in 1:18)
  final[,i] <- factor(final[,i], levels=c('Strongly Disagree','Disagree','Somewhat Agree','Agree','Strongly Agree'))




```

Finally printing the graph into a pdf 

```{r}
f <- likert(final, grouping=row.names(final))
pdf(file='facevapharm.pdf',width=8.5, height=13, title='Faculty Evaluation',onefile=T)
plot(f)

names = unique(al[,5])
for (k in names[1:4])
  {
  k = as.character(k)
  fm <- final[k == row.names(final) ,]
  #row.names(fm) <- row.names(final[names[k] == row.names(final) ,])
  
  f <- likert(fm)
  print(plot(f))
  }

dev.off()

```


As some additions the question descriptions can be added and percents at the center may be dropped in plotting options.





