FACULTY EVALUATION ANALYSIS USING LIKERT SCALE
========================================================

First we transpose the data from column 9 to 109 
library(likert)
library(reshape)
```{r}

fil <- choose.files()
if(!grep('.csv$',fil))   return()
al <- read.csv(fil,header=T)
resp <- t(al[,9:108])
colnames(resp) <- t(al[,8]) 

```

We trim unused rows at the last filled with the NAs and add faculty member names so many
times as many are observations for each
```{r}
resp<- resp[rowSums(resp,na.rm=T)!=0,]
obs<-c()
for (i in seq(1,ncol(resp),18)) 
  obs <- append(obs,sum(rowSums(resp[,i:c(i+17)],na.rm=T)!=0))

#splitting each succesive 18 columns and binding them  down in to rows
totalobs = max(obs*length(obs))
maxobs = max(obs)
final<- c(NULL)
i=1
for (j in seq(1, totalobs, maxobs) )
  {
  final <- rbind(final, resp[1:maxobs,i:c(i+17)])
  i=i+18
}

## adding faculty member names along with course names as row names
row.names(final) <- rep(unique(al[,5]), each=maxobs)

## cleaning rows with full NAs
final <- final [rowSums(final,na.rm=T)!=0,]
write.csv(final, 'FacEvalforlikert.csv')



```

Upto here we get clean data if we also want to include department names then we could 
add a column in the beggining and add just like we added faculty member names

Next we replace values 1 to 5 with the character equivalents and converting them into factors CAUTION: HERE FACTORS SHOULD BE DEFINED EXPLICITLY
```{r}
final <- as.data.frame(final)
final[final==5]<-'Strongly Agree'
final[final==4]<-'Agree'
final[final==3] <- 'Somewhat Agree'
final[final==2] <- 'Disagree'
final[final==1] <- 'Strongly Disagree'

for (i in 1:18)
  final[,i] <- factor(final[,i], levels=c('Strongly Disagree','Disagree','Somewhat Agree','Agree','Strongly Agree'))

colnames(final) <- al[1:18,7]


```

Finally printing the graph into a pdf 

```{r}
n <- strsplit(fil,'[\\.]')
n <- n[[1]][length(n[[1]])-1]  #get the filename

f <- likert(final, grouping=c(row.names(final)))
#xtable(f)
pdf(file=paste(n,'.pdf',''),width=8.5, height=7*length(obs))
plot(f) + ggtitle(n)
dev.off()

```


As some additions the question descriptions can be added and percents at the center may be dropped in plotting options.








