Loans Issued through a Lending Club
========================================================
library(knitr) # required for generating R Markdown like this one

INTRODUCTION:
The Loans Data is collected from https://spark-public.s3.amazonaws.com/dataanalysis/loansData.rda, which includes 2500 peer to peer loans issued through the Lending Club 
https://www.lendingclub.com/home.action

library(psych)
library(multilevel)
library(car)
library(ggplot2)
library(reshape)

```{r}


#download.file(url='https://spark-public.s3.amazonaws.com/dataanalysis/loansData.rda',destfile='loans.rda')
load('../Assignment 1/data/loansData.rda')
summary(loansData)

```

First we check if there are any NA values in file and tidying the data
```{r}
loansData$Employment.Length[loansData$Employment.Length == 'n/a'] <- NA
loansData$Employment.Length[loansData$Employment.Length == '< 1 year'] <- 0.5
loansData$Employment.Length <- as.numeric(gsub(' years$| year$|\\+','',loansData$Employment.Length))
attach(loansData)
sum(is.na(loansData))
loans <- na.omit(loansData)
attach(loans)
summary(loans)
```

We now have 2298 obs. of 14 variables of data without NA's; we observe that Interest Rate and Debt to Income Ratio is in % so we clean that. Also we fine tuned the FICO Range column to make it easier to calculate means of the ranges:
```{r}
loans[,3] <- as.numeric(gsub('%','',loans[,3]))
loans[,6] <- as.numeric(gsub('%','',loans[,6]))
library(reshape)
loans <- transform(loans, FICO.Range = colsplit(FICO.Range, split='-', names=c('low','high')))
attach(loans)

```

THE EXPLORATORY ANALYSIS

```{r}
hist(rowMeans(FICO.Range), breaks=20)
summary(rowMeans(FICO.Range))
mean(rowMeans(FICO.Range))
boxplot(Monthly.Income ~ State)
loans <- loans[Monthly.Income < 30000,]
```

The histogram of means of FICO Ranges is heavily right skewed with mean 707.74 and overall mean ranges from 642 to 832 dispersed over a 190 interval. 

CONFOUNDERS IN Monthly Income:
There are obvious confounders in Monthly specially in New York.

Next we try to find relationship between Interest Rate and Mean FICO Ranges

```{r}
cor.test( rowMeans(FICO.Range), Interest.Rate)

```

We get correlation of -0.71 with p-value < 2.2e-16. Hence Interest Rate is negatively associated with the FICO Range. Greater the FICO Score the borrower is likely to get the loan at a lower interest rate overall in all states.

for this lets first make a basic model:

```{r}
model1 <- lm(Interest.Rate ~ rowMeans(FICO.Range))
summary(model1)
```
R-squared is 50% which is the variance in Interest Rate defined by the FICO Range.

Lets see if any state has any reservation on that trend:

```{r}
model2 <- lm(Interest.Rate ~ rowMeans(FICO.Range)+State)
summary(model2)
```

We see that it is consistent accross almost all states considering the significant p-values.



```{r}
model3 <- lm(Interest.Rate ~ rowMeans(FICO.Range)+Monthly.Income+Revolving.CREDIT.Balance+Inquiries.in.the.Last.6.Months)
summary(model3)
```
Not much difference yet

```{r}
cor (Interest.Rate, Revolving.CREDIT.Balance)


```

.06

```{r}
#library(ggplot2)
model4 <- lm(Interest.Rate ~  log10(Amount.Funded.By.Investors+1) + rowMeans(FICO.Range) )
summary(model4)
ggplot(loans, aes(x=rowMeans(FICO.Range),y=Interest.Rate))+geom_smooth(method = 'lm')+geom_point(aes(color=Amount.Funded.By.Investors))+ggtitle('Fig 2')
```
Greater the amount funded greater the interest t=32.65 for Amount by Investors

```{r}
model5 <- lm (Interest.Rate ~ rowMeans(FICO.Range) +Employment.Length)
summary(model5)
confint(model5)
```
A very weak correlation between employment length and Interest Rate
OUR FINAL MODEL IS 
```{r}
model6a <- lm(Interest.Rate ~ rowMeans(FICO.Range) + Amount.Funded.By.Investors+Loan.Length)

modi <- influence.measures(model6)
inf <- which (apply(modi$is.inf,1,any))
a <- 1:nrow(loans)
model6 <- lm(Interest.Rate ~ rowMeans(FICO.Range) + Amount.Funded.By.Investors+Loan.Length, subset=(a[-inf]))

plot(rowMeans(FICO.Range), model6$residuals, col=c(as.numeric(Loan.Length)*100), pch=19, main='Fig 3: Colored Residuals factored on Loan Lengths',xlab='FICO Score',ylab='Residuals')
plot(model6$fitted, model6$residuals, col=c(as.numeric(Loan.Length)*100), main='Fig 4: Colored Model factored on Loan Lengths',xlab='Fitted Values',ylab='Residuals',pch=19)

plot(rowMeans(FICO.Range), model6$residuals, cex=Amount.Funded.By.Investors/max(Amount.Funded.By.Investors), pch=19,main='Sized Residual ratios on Amount Funded by Investor' )

plot(model6$fitted, model6$residuals, cex=Amount.Funded.By.Investors/max(Amount.Funded.By.Investors), main='Sized point ratios on Amount Funded by Investors',pch=19)

hist(FICO.Range,main='FICO Range Histogram')

summary(model6)

```




```{r}


l[,c(4)] <- as.numeric(as.factor(loans[,c(4)]))
l[,c(5)] <- as.numeric(as.factor(loans[,c(5)]))
l[,c(7)] <- as.numeric(as.factor(loans[,c(7)]))
l[,c(8)] <- as.numeric(as.factor(loans[,c(8)]))
l$meanFICO <- rowMeans(FICO.Range)
library(ellipse)
ctab=cor(l[,-10])
windows()
colorfun <- colorRamp(c('yellow','skyblue','red'), space="Lab")
plotcorr(ctab, col=rgb(colorfun((ctab+1)/2), maxColorValue=255))
title('Fig 1: Correlation Matrix')

model7 <- lm(log10(Monthly.Income) ~ Home.Ownership+Employment.Length+Loan.Purpose)
summary(model7)
```

MEDIATIONAL ANALYSIS
```{r}
library(multilevel)
attach(loans)
model.sobel <- sobel( Inquiries.in.the.Last.6.Months ,rowMeans(FICO.Range), Interest.Rate)
model.sobel

model.sobel2 <- sobel( as.numeric(as.factor(Home.Ownership)) ,rowMeans(FICO.Range), Interest.Rate)
model.sobel2


```

```{r}
model.sobel3 <- sobel(pred=log10(Amount.Requested),med=as.numeric(as.factor(Loan.Length)),out=Interest.Rate)
model.sobel3


```

MODERATIONAL ANALYSIS
```{r}

model8 <- lm(Amount.Funded.By.Investors ~ rowMeans(FICO.Range)+ Interest.Rate+ (Interest.Rate*rowMeans(FICO.Range)))
model81 <- lm(Amount.Funded.By.Investors ~ rowMeans(FICO.Range)+Interest.Rate)
anova(model81, model8)

#summary(model8)

```

Looking at modelling tree we can see there is some complex machanism going on in selection of interest rates; and the most effect comes from FICO range.Then for different fico ranges the decision is furthered on amount funded, yet the main effect is that of FICO

```{r}
pdf(height=13, width=13)
mod0 <- tree(Interest.Rate ~ rowMeans(FICO.Range)+Amount.Funded.By.Investors+Revolving.CREDIT.Balance+Monthly.Income+Home.Ownership)
dev.off()
```





Let's see effect of FICO on amount funded is nested on  Interest rate 

```{r}
model9 <- lm(Amount.Funded.By.Investors ~  rowMeans(FICO.Range)^2 / Interest.Rate)
summary(model9)

```
OR THE INTERACTION
```{r}
model91 <- lm(Amount.Funded.By.Investors ~  rowMeans(FICO.Range) : Interest.Rate)
summary(model91)
```

THOSE WHO HAVE HOMES HAVE RELATIVELY LOWER INCOME AND THEY ARE CORRELATED WITH SOME STATES








```

















attach(loansData)
model1 <- lm(Interest.Rate ~ Debt.To.Income.Ratio)
summary(model1) # Only 3% variance is explained by this model
summary(Employment.Length)
tapply(Interest.Rate, Employment.Length, sum) # No relationship
model2 <- lm(Interest.Rate ~ Inquiries.in.the.Last.6.Months)
summary(model2) # Only 2.7 variance explained
model3 <- lm(Interest.Rate ~ Revolving.CREDIT.Balance)
summary(model3) # no luck
model4 <- lm(Interest.Rate ~ Open.CREDIT.Lines)
summary(model4) # no luck
## LETS PLAY WITH FICO RANGE
library(reshape)
loansData <- transform(loansData, FICO.Range = colsplit(FICO.Range, split='-', names=c('low','high')))
attach(loansData)
loansData$meanFICO <- rowMeans(FICO.Range)
attach(loansData)
head(loansData)
model5<- lm (Interest.Rate ~ meanFICO)
summary(model5)    # WOW! 50% variance is defined by the model and it is significant 
## LETS DO THE MEDIATION ANALYSIS BY SOBEL FUNCTION
library(multilevel)
model.sobel <- sobel(meanFICO, Inquiries.in.the.Last.6.Months , Interest.Rate)
model.sobel # z value is -3.88 which is high enough, so inquiries mediates the effect statistically negatively
model.sobel2 <- sobel( Inquiries.in.the.Last.6.Months ,meanFICO, Interest.Rate)
model.sobel2 # z value is 4.6 which is higher enough, so meanFICO mediates the effect


